---
title: "Train GLMNET"
author: "RFB"
date: "28 1 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(glmnet)
library(glmnetUtils)
library(mice)
library(reshape2)
library(parallel)
library(foreach)
library(doParallel)
library(ggplot2)
library(plyr)
library(tidyverse)
library(matrixStats)

#setwd("~/Rprojects/dream_beat_aml/Scripts/")
```

Build X matrix
```{r}
source("aux_functions.R")
X <- buildX("../ExternalData/")
```

Load AUC data, reshape and impute missing
```{r}
aucs <- read.csv("../ExternalData/aucs.csv", stringsAsFactors = F)
aucs_wide <- dcast(aucs, lab_id~inhibitor, value.var = "auc")
aucs_wide_notNA <- mice(data.frame(aucs_wide[,-1]), meth = "cart", m = 1)
y <- complete(aucs_wide_notNA)
colnames(y) <- colnames(aucs_wide[,-1])
```

## Multivariate model
```{r}
### Submitted to validation
if(!file.exists("../GeneratedData/dose_response_model_mv_en.RData")){
  ## Fit glmnet
  cl <- makeForkCluster(5)
  
  fit_mv <- cva.glmnet(x = as.matrix(X),
                       y = as.matrix(y),
                       family = "mgaussian",
                       outerParallel = cl,
                       #parallel = T,
                       alpha = seq(0.5,0.95, by = 0.05))
  stopCluster()
  
  bestFit <- which.min(sapply(fit_mv$modlist, function(x) min(x$cvm)))
  dose_response_model <- fit_mv$modlist[[bestFit]]
  save(dose_response_model, file = "../GeneratedData/dose_response_model_mv_en.RData")
} else{
  load("../GeneratedData/dose_response_model_mv_en.RData")
}
```

## Single variate models
```{r}
if(!file.exists("../GeneratedData/dose_response_model_en.RData")){
  # Set up parallel computations
  registerDoParallel(cores = 5)
  fit <- foreach(i=1:ncol(y),
                 .errorhandling = "pass",
                 .packages = c("glmnetUtils")) %dopar% {
                   cva.glmnet(x = as.matrix(X),
                              y = as.matrix(y[,i]),
                              family = "gaussian",
                              parallel = T)
                 }
  findMin <- function(x){
    a <- which.min(sapply(x$modlist, function(x) min(x$cvm)))
    x$modlist[[a]]
  }
  
  dose_response_models <- lapply(fit, findMin)
  names(dose_response_models) <- colnames(aucs_wide)[-1]
  save(dose_response_models, file = "../GeneratedData/dose_response_model_en.RData")
} else{
  load("../GeneratedData/dose_response_model_en.RData")
}
```


# Stats for presentation
## Univariate
Total MSE
```{r}
sum(sapply(dose_response_models, function(x) x$cvm[x$lambda == x$lambda.min]))
```


plot figure
```{r, fig.height=7, fig.width=8}
plot(dose_response_models[[14]])
title("Bortezomib (Velcade)", line = -2)
```


Predict data
```{r}
pred_fit <- list()
for(drug in unique(aucs$inhibitor)){
  fit <- dose_response_models[[drug]]
  pred_fit[[drug]] <- predict(fit, 
                              newx = as.matrix(X[,row.names(coef(fit))[-1]]), 
                              s = "lambda.min")
}
```

Calculate and plot results in training data
```{r}
train_cor <- c()
for(drug in unique(aucs$inhibitor)){
  temp <- merge(data.frame(pred_fit[[drug]]), 
                subset(aucs, inhibitor == drug),
                by.x = 0, by.y = "lab_id")
  train_cor[drug] <- cor(temp$X1, temp$auc, method = "spearman")
}

plot(train_cor, main = "Correlation of true vs pred in training", ylab = "Cor")
abline(h = mean(train_cor, na.rm = T), col = "red")
```

## Multivariate
MSE
```{r}
dose_response_model$cvm[dose_response_model$lambda == dose_response_model$lambda.min]
```

Plot figure
```{r, fig.height=7, fig.width=8}
plot(dose_response_model)
title("Multivariate model", line = -2)
```


Predict from model
```{r}
pred_fit_mv <- predict(dose_response_model,
                       newx = as.matrix(X),
                       s = "lambda.min")

pred_fit_mv_long <- melt(pred_fit_mv[,,1])
names(pred_fit_mv_long) <- c("lab_id", "inhibitor", "auc")
```

Calculate and plot results in training data
```{r}
train_cor_mv <- c()
for(drug in unique(aucs$inhibitor)){
  train_cor_mv[drug] <- cor(aucs_wide[,drug], pred_fit_mv[,drug,1], method = "spearman", use = "pairwise.complete.obs")
}

plot(train_cor_mv, main = "Correlation of true vs pred in training (mv model)", ylab = "Cor")
abline(h = mean(train_cor_mv, na.rm = T), col = "red")
```

## Compare uni vs multi model
```{r}
auc_cor <- cor(aucs_wide[,-1], use = "pairwise.complete.obs", method = "pearson")
auc_cor2 <- cor(aucs_wide[,-1], use = "pairwise.complete.obs", method = "spearman")


diag(auc_cor) <- NA

plotData <- data.frame("uni" = train_cor[order(names(train_cor))],
                       "multi" = train_cor_mv[order(names(train_cor_mv))],
                       "mean_drug_cor_pearson" = colMeans(auc_cor, na.rm = T),
                       "median_drug_cor_pearson" = colMedians(auc_cor, na.rm = T),
                       "mean_drug_cor_spearman" = colMeans(auc_cor2, na.rm = T),
                       "median_drug_cor_spearman" = colMedians(auc_cor2, na.rm = T),
                       "pct_miss" = colSums(is.na(aucs_wide[,-1]))/nrow(aucs_wide))
plotData$uni[is.na(plotData$uni)] <- 0

ggplot(plotData, aes(x = uni, y = multi, col = mean_drug_cor_pearson)) +
  geom_point() +
  scale_color_viridis_c()

ggplot(plotData, aes(x = uni, y = multi, col = pct_miss)) +
  geom_point() +
  scale_color_viridis_c()

mean_diff <- mean(plotData$multi - plotData$uni, na.rm = T)

ggplot(plotData, aes(x =1:nrow(plotData) , y = (multi-uni), col = mean_drug_cor_pearson)) +
  geom_point() +
  xlab("") + 
  ylab("Difference (multi - uni)") +
  scale_color_viridis_c(name = "Mean correlation") +
  geom_hline(yintercept = mean_diff) +
  ggtitle("Difference in training data pearson correlation of Multi vs Uni")
```


```{r}
plotData %>% rownames_to_column(var = "drug") %>% 
write_csv("../GeneratedData/multi_uni_plot_data.csv")
```


Find drugs with high performance in multi model
```{r}
knitr::kable(plotData[(plotData$multi - plotData$uni) > 0.3,])
```

Check average correlation vs performance of multi model
```{r}
plotData %>% 
  mutate(Multi_model = ifelse(multi > uni, "Better", "Worse")) %>% 
  ggplot(aes(x = Multi_model, y = mean_drug_cor_pearson)) + geom_boxplot() +
  ggpubr::stat_compare_means()
```


Check COEFS
```{r}
g <- coefficients(dose_response_model)
drugname <- names(g)

g <- do.call(cbind,g)
g <- g[rowSums(g) != 0, ]
g <- as.matrix(g)
colnames(g) <- drugname

write.csv(as.matrix(g), file = "../GeneratedData/dose_response_multi_coef.csv")
```



```{r}
tail(row.names(g), n = 20)

table(X[, "dxAtInclusionMYELODYSPLASTIC SYNDROMES"], X[, "dxAtSpecimenAcquisitionMYELODYSPLASTIC SYNDROMES"])

plot(g["dxAtInclusionMYELODYSPLASTIC SYNDROMES", ] - g["dxAtSpecimenAcquisitionMYELODYSPLASTIC SYNDROMES", ])
plot(g["dxAtInclusionMYELODYSPLASTIC SYNDROMES", ] + g["dxAtSpecimenAcquisitionMYELODYSPLASTIC SYNDROMES", ])
```

```{r}
plot(g["dna_FABP2",])
plot(g["dna_HECW2",])
plot(g["dna_LRP8",])
plot(g["dna_PKD1L2",])
plot(g["dna_PLRG1",])
plot(g["dna_TMC3",])
```

